Spatiotemporal clusters in the CSD-CM representation

### 2019-09-04 Reportes/clusteringkarel.tex
mas figuras en reporte
y ahora tablas

@@

\caption{Same data as in fig. \ref{puntostodos} but colorized using
the temporal appearence. The colorbar is in miliseconds.
}\label{puntosconcolor}

@@

Clustering algorithms vary not only in implementation, but in design criteria.
Thus they produce different results. We had to test a few of them. Our choice at the
end was the so called DBSCAN (density based spatial clustering of applications with noise).
This algorithm is conceptually clean and can produce heterogeneous clusters, in contrast
to others such as HDC ( Hierarchical Data clustering). The DBSCAN clustering depends
on two parameters: a radio of neighborhood and a minimal number of neighbors. Each
point that is deemed to be part of a certain cluster has to have that number of neighbors
that are also part of the same set inside an sphere of the chosen radius. These
parameters can be chosen on physiological grounds although we can device a more
abstract way of optimizing them.

We decided to use a certain ``aesthetic'' criterium for choosing the optimal parameters
for the DBSCAN algorythm.  It is its accordance with Zipf's Law.  This could post out
the accordance with the allometric inverse algebraic laws that are observed in
anatomical studies, but it is, in the end, a somewhat arbitrary criterium. Even thou, a good
fit with Zipfs Law could be an indictor of a natural clustering of data points.
A problem that appeard using this criterium is that it is such a pervarsive
behaviour, that it gives a good fit to allmost any set of reazonable parameters
that one encounters. So we had to come up with criteria for what these ``reasonable
parameters'' may be.  We settled that we shouldn't have to many unasigned points, and
that no cluster should be too big, or that there were to many clusters.

>>>>

\author{K. Zapfe}
\title{Spatiotemporal clusters in the CSD-CM representation}

@@

\caption{Same data as in fig. \ref{puntostodos} but colorized using
the temporal appearence. The colorbar is in miliseconds. We can observe
that there is near simultaneus activation of the ``Sources Layers'' in CA3,
between other visual groupings. 
}\label{puntosconcolor}

@@

Clustering algorithms vary not only in implementation, but in design criteria.
Thus they produce different results. We had to test a few of them. Our choice at the
end was the so called DBSCAN (Density Based Spatial Clustering of Applications with Noise).
This algorithm is conceptually clean and can produce heterogeneous clusters, in contrast
to others such as HDC ( Hierarchical Data clustering). The DBSCAN clustering depends
on two parameters: a radio of neighborhood and a minimal number of neighbors. Each
point that is deemed to be part of a certain cluster has to have that number of neighbors
that are also part of the same set inside an sphere of the chosen radius. These
parameters can be chosen on physiological grounds although we can device a more
abstract way of optimizing them. An example of the results of this
algorythm is shown on fig. \ref{puntoscluster}.
A strict choice of parameters (small radius and many neighbors) produces
very dense clusters and a lot of unasigned points. A laxer choice (a big radius
with one to three neighboors) produces few clusters, with sparcely populated
areas,  and fewer unasigned points. A quick testing of the algorythm 
shows that we have to explore the validity of clusters in the range of
$r \in [0.75, 2 ]$ ( electrode lengths) and $ n \in \lbrace 2,3,4,5 \rbrace$.

\begin{figure}[h]
\centering
\begin{subfigure}{0.40\textwidth}
\includegraphics[width=\textwidth]{dbscannegejem01.png}
\caption{Sinks}
\end{subfigure}
\begin{subfigure}{0.40\textwidth}
\includegraphics[width=\textwidth]{dbscanposejem01.png}
\caption{Sources}
\end{subfigure}

\caption{Same data as in previous figures, but the colorization is
  from a realization of DBSCAN clustering. The parameters used were
  $r=1.75$ a.u. for the neighborhood radius and 3 neighbors for each point.
  Those parameters give around 300 clusters, many of them with less than
  10 points. A good choice of parameters requires balancing
  the number of clusters and the number of unasigned points. 
}\label{puntoscluster}

We decided to use a certain ``aesthetic'' criterium for choosing the optimal parameters
for the DBSCAN algorythm.  It is its accordance with Zipf's Law.  This could post out
the accordance with the allometric inverse algebraic laws that are observed in
anatomical studies, but it is, in the end, a somewhat arbitrary criterium.
Nevertheless, a good
fit with Zipfs Law could be an indictor of a natural clustering of data points.
A problem that appeard using this criterium is that it is such a pervarsive
behaviour of ranked data
that it gives a good fit to allmost any set of reazonable parameters
that one encounters. So we had to come up with criteria for what these ``reasonable
parameters'' may be.  We settled that we shouldn't have to many unasigned points, and
that no cluster should be too big, or that there were to many clusters.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{zipfexample.png}
  \caption{An instance of Zipf's Law, with DBSCAN parameters
    $r=1.5$ a.u. and $n=4$. The agreement with Zipf's
    gives an exponent of $-1.007$ against the ideal $-1$ and
    an $R^2$ of $0.986$.}
\end{figure}

\begin{tabular}{|r|r|r|r|r|r|}
 radius &  n  & $R^2$ & u & m & ng  \\ \hline
  1 &  2 &  0.978 & 606 & 606 & 1285 \\
  1 &  3 &  0.984 & 1346 & 1346 & 915 \\
  1 &  4 &  0.974 & 2300 & 2300 & 658 \\
  1 &  5 &  0.951 & 3193 & 3193 & 493 \\
  1 &  6 &  0.918 & 4023 & 4023 & 375 \\
  1.5 &  2 &  0.987 & 276 & 1007 & 691 \\
  1.5 &  3 &  0.987 & 660 & 1007 & 499 \\
  1.5 &  4 &  0.985 & 1060 & 1060 & 391 \\
  1.5 &  5 &  0.986 & 1523 & 1523 & 315 \\
  1.5 &  6 &  0.99 & 2040 & 2040 & 258 \\
  2 &  2 &  0.95 & 156 & 5355 & 343 \\
  2 &  3 &  0.928 & 382 & 5355 & 230 \\
  2 &  4 &  0.912 & 566 & 5349 & 174 \\
  2 &  5 &  0.906 & 777 & 5308 & 132 \\
  2 &  6 &  0.915 & 976 & 3734 & 112 \\
  2.5 &  2 &  0.911 & 113 & 6144 & 225 \\
  2.5 &  3 &  0.875 & 243 & 6144 & 160 \\
  2.5 &  4 &  0.849 & 349 & 6142 & 127 \\
  2.5 &  5 &  0.828 & 475 & 6127 & 100 \\
  2.5 &  6 &  0.849 & 602 & 6008 & 81 \\
  3 &  2 &  0.909 & 89 & 7408 & 163 \\
  3 &  3 &  0.874 & 187 & 7408 & 114 \\
  3 &  4 &  0.847 & 267 & 7404 & 89 \\
  3 &  5 &  0.829 & 343 & 7402 & 71 \\
  3 &  6 &  0.817 & 418 & 7400 & 58 
\end{tabular}
